 ---
title: "Old precision and recall methods"
output: html_document
author: Max Dippel
date: "2023-09-12"
---

This rmd contains all of the old and dysfunctional ways to calculate precision and recall that we are no longer using

Precision is the fraction of relevant instances divided by all retrieved instances. This is also called the positive predictive value. 
Precision = True Positives / (True Positives + False Positives)

Recall is the fraction of all retrieved instances divided by all the retrieved instances. This is also called the sensitivity. 

Precision = True Positives / (True Positives + False Negatives)

Loading the MPRA data (this is the August 25th primary T cell using the T cell filter)
```{r, eval=FALSE, echo=FALSE}
mpra.dir <- "/nfs/jray/screens/Mouri_et_al_MPRA/Mouri_et_al_replication/Mouri_replication_rmds/20230825_stimulated_primary_Tcell_glm_replication/john_filter/data/20230825_mpra_merge_Tcell_glm_john.txt"
mpra <- read.table(mpra.dir, sep="\t", header=T)
```

This runs a loop to find the max fold changes.
```{r, eval=FALSE, echo=FALSE}
#For each MPRA variant, find the fold change of the allele with the largest absolute fold change
# This creates a column full of NAs for the loop to fill
mpra$max_abs_fc <- NA
# This loop calculates the column of max_abs_fc. It takes the two alleles A and B and calculates which one's absolute value is larger. That larger value is the value of the new column (can be positive or negative). 
for(i in 1:nrow(mpra)){
  if(abs(mpra[i,]$A.log2FC)>abs(mpra[i,]$B.log2FC)){
    mpra[i,]$max_abs_fc<-mpra[i,]$A.log2FC
  }else{
    mpra[i,]$max_abs_fc<-mpra[i,]$B.log2FC
  }
}
```

This runs a loop to find the max fold change p values.
```{r, eval=FALSE, echo=FALSE}
#For each MPRA variant, find the fold change of the allele with the largest absolute fold change
# This creates a column full of NAs for the loop to fill
mpra$max.logPadj_BH<-NA
# This loop calculates the column of max_abs_fc. It takes the two alleles A and B and calculates which one's absolute value is larger. That larger value is the value of the new column (can be positive or negative). 
for(i in 1:nrow(mpra)){
  if(abs(mpra[i,]$A.logPadj_BH)>abs(mpra[i,]$B.logPadj_BH)){
    mpra[i,]$max.logPadj_BH<-mpra[i,]$A.logPadj_BH
  }else{
    mpra[i,]$max.logPadj_BH<-mpra[i,]$B.logPadj_BH
  }
}
```

First attempt at precision and recall calculations without that many fancy tricks 
```{r, eval=FALSE, echo=FALSE}
# creating empty dataframe to populate with ones and zeros based on fc and p values 
dat.precision<- data.frame(mpra$max.logPadj_BH,abs(mpra$max_abs_fc))
names(dat.precision) <- c("max.logPadj_BH","max_abs_fc")
dat.precision$max.logPadj_BH_0.25=0 # max.logPadj_BH greater than or equal to 0.25
dat.precision$max.logPadj_BH_0.5=0 # max.logPadj_BH greater than or equal to 0.5
dat.precision$max.logPadj_BH_1=0 # max.logPadj_BH greater than or equal to 1
dat.precision$max.logPadj_BH_2=0 # max.logPadj_BH greater than or equal to 2
dat.precision$max_abs_fc_0.25=0 # max_abs_fc greater than or equal to 0.25
dat.precision$max_abs_fc_0.5=0 # max_abs_fc greater than or equal to 0.5
dat.precision$max_abs_fc_1=0 # max_abs_fc greater than or equal to 1
dat.precision$max_abs_fc_2=0 # max_abs_fc greater than or equal to 2

# Big for loop to populate 
for(i in 1:nrow(dat.precision)){
   if(dat.precision[i,]$max.logPadj_BH>=0.25){ 
    dat.precision[i,]$max.logPadj_BH_0.25=1
   } # If max.logPadj_BH greater than or equal to 0.25 then put a 1 in the max.logPadj_BH_0.25 column
   if(dat.precision[i,]$max.logPadj_BH>=0.5){
    dat.precision[i,]$max.logPadj_BH_0.5=1
   } # If max.logPadj_BH greater than or equal to 0.5 then put a 1 in the max.logPadj_BH_0.5 column
   if(dat.precision[i,]$max.logPadj_BH>=1){
    dat.precision[i,]$max.logPadj_BH_1=1
  } # If max.logPadj_BH greater than or equal to 1 then put a 1 in the max.logPadj_BH_1 column
  if(dat.precision[i,]$max.logPadj_BH>=2){
    dat.precision[i,]$max.logPadj_BH_2=1
  } # If max.logPadj_BH greater than or equal to 2 then put a 1 in the max.logPadj_BH_2 column
  if(dat.precision[i,]$max_abs_fc>=0.25){
    dat.precision[i,]$max_abs_fc_0.25=1
   } # If max_abs_fc greater than or equal to 0.25 then put a 1 in the max_abs_fc_0.25 column
  if(dat.precision[i,]$max_abs_fc>=0.5){
    dat.precision[i,]$max_abs_fc_0.5=1
  } # If max_abs_fc greater than or equal to 0.5 then put a 1 in the max_abs_fc_0.5 column
  if(dat.precision[i,]$max_abs_fc>=1){
    dat.precision[i,]$max_abs_fc_1=1
   } # If max_abs_fc greater than or equal to 1 then put a 1 in the max_abs_fc_1 column
   if(dat.precision[i,]$max_abs_fc>=2){
    dat.precision[i,]$max_abs_fc_2=1
   } # If max_abs_fc greater than or equal to 2 then put a 1 in the max_abs_fc_2 column
}
```

Calculating the precision values and then creating a heat map
```{r, eval=FALSE, echo=FALSE}
df <- data.frame(cbind(
c(Metrics::precision(dat.precision$max.logPadj_BH_0.25, dat.precision$max_abs_fc_0.25), 
Metrics::precision(dat.precision$max.logPadj_BH_0.5, dat.precision$max_abs_fc_0.25),
Metrics::precision(dat.precision$max.logPadj_BH_1, dat.precision$max_abs_fc_0.25),
Metrics::precision(dat.precision$max.logPadj_BH_2, dat.precision$max_abs_fc_0.25)),

c(Metrics::precision(dat.precision$max.logPadj_BH_0.25, dat.precision$max_abs_fc_0.5),
Metrics::precision(dat.precision$max.logPadj_BH_0.5, dat.precision$max_abs_fc_0.5),
Metrics::precision(dat.precision$max.logPadj_BH_1, dat.precision$max_abs_fc_0.5),
Metrics::precision(dat.precision$max.logPadj_BH_2, dat.precision$max_abs_fc_0.5)),

c(Metrics::precision(dat.precision$max.logPadj_BH_0.25, dat.precision$max_abs_fc_1),
Metrics::precision(dat.precision$max.logPadj_BH_0.5, dat.precision$max_abs_fc_1),
Metrics::precision(dat.precision$max.logPadj_BH_1, dat.precision$max_abs_fc_1),
Metrics::precision(dat.precision$max.logPadj_BH_2, dat.precision$max_abs_fc_1)),

c(Metrics::precision(dat.precision$max.logPadj_BH_0.25, dat.precision$max_abs_fc_2),
Metrics::precision(dat.precision$max.logPadj_BH_0.5, dat.precision$max_abs_fc_2),
Metrics::precision(dat.precision$max.logPadj_BH_1, dat.precision$max_abs_fc_2),
Metrics::precision(dat.precision$max.logPadj_BH_2, dat.precision$max_abs_fc_2))
))

colnames(df) <- c("FC 0.25","FC 0.5","FC 1", "FC 2")
rownames(df) <- c("Padj_BH 0.25", "Padj_BH 0.5", "Padj_BH 1", "Padj_BH 2")

df2 <- df %>%
  rownames_to_column() %>%
  gather(colname, value, -rowname)
head(df2)

ggplot(df2, aes(x = rowname, y = colname, fill = value)) +
  geom_tile() + geom_text(aes(label = round(value, digit=2))) + xlab("Max log p-value Benjamini-Hochberg adjustment") + ylab("Max absolute value fold change") + ggtitle("Precision heatmap for Fold Change") + 
  scale_fill_gradient(low = "gold", high = "firebrick2") +
  coord_fixed()
```

Loop to give zeros and ones based on fc values
```{r, eval=FALSE, echo=FALSE}
dat.precision<- data.frame(abs(mpra$max_abs_fc)) # create a data frame with max.logPadj_BH and max_abs_fc 
a[1:20] <- data.frame(rep.int(0, 13845)) # create a data frame of zeros
dat.precision2 <- data.frame(dat.precision,a) # add the max.logPadj_BH and max_abs_fc columns
names(dat.precision2) <- c("fc",paste0("fc_", seq(0.1, 2, by = 0.1))) # give names to the columns
dat.precision2 <- dat.precision2[,-22]

dat.precision2 <- dat.precision2 |>
  mutate(row = row_number()) |> # not necessary if `fc` values guaranteed unique
  pivot_longer(-c(row, fc)) |>
  mutate(value = 1 * (parse_number(name) <= fc)) |>
  pivot_wider(names_from = name, values_from = value)
```

Loop to give zeros and ones based on padj values
```{r, eval=FALSE, echo=FALSE}
dat.precision3<- data.frame(mpra$max.logPadj_BH) # create a data frame with max.logPadj_BH and max_abs_fc 
a <- c(0,0,0)
a[1:20] <- data.frame(rep.int(0, 13845)) # create b data frame of zeros
dat.precision3 <- data.frame(dat.precision3,a) # add the max.logPadj_BH and max_abs_fc columns
names(dat.precision3) <- c("padj",paste0("padj_", seq(0.1, 2, by = 0.1))) # give names to the columns

dat.precision3 <- dat.precision3 |>
  mutate(row = row_number()) |> # not necessary if `padj` values guaranteed unique
  pivot_longer(-c(row, padj)) |>
  mutate(value = 1 * (parse_number(name) <= padj)) |>
  pivot_wider(names_from = name, values_from = value)
```

Merge the two data frames
```{r, eval=FALSE, echo=FALSE}
dat.precision4 <- merge(dat.precision2, dat.precision3, by="row")
dat.precision4 <- dat.precision4[,-1]
```

Calculating the precision values and then creating a heat map
```{r, eval=FALSE, echo=FALSE}
df <- data.frame(cbind(
c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.1), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.1), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.1)),

c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.2), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.2), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.2)),

c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.3), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.3), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.3)),

c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.4), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.4), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.4)),

c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.5), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.5), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.5))
))

colnames(df) <- c("FC 0.1","FC 0.2","FC 0.3", "FC 0.4", "FC 0.5")
rownames(df) <- c("Padj_BH 0.1", "Padj_BH 0.2", "Padj_BH 0.3", "Padj_BH 0.4", "Padj_BH 0.5", "Padj_BH 0.6", "Padj_BH 0.7", "Padj_BH 0.8", "Padj_BH 0.9", "Padj_BH 1")

df2 <- df %>%
  rownames_to_column() %>%
  gather(colname, value, -rowname)
head(df2)

ggplot(df2, aes(x = rowname, y = colname, fill = value)) +
  geom_tile() + geom_text(aes(label = round(value, digit=2))) + xlab("Max log p-value Benjamini-Hochberg adjustment") + ylab("Max absolute value fold change") + ggtitle("Precision heatmap for Fold Change") + 
  scale_fill_gradient(low = "gold", high = "firebrick2") +
  coord_fixed()

```


###########################################

Specific for high pics values

###########################################

Loading the MPRA data (this is the August 25th primary T cell using the T cell filter)
```{r, eval=FALSE, echo=FALSE}
mpra.dir <- "/nfs/jray/screens/Mouri_et_al_MPRA/Mouri_et_al_replication/Mouri_replication_rmds/20230825_stimulated_primary_Tcell_glm_replication/john_filter/data/20230825_mpra_merge_Tcell_glm_john.txt"
mpra <- read.table(mpra.dir, sep="\t", header=T)
```

This runs a loop to find the max fold changes.
```{r, eval=FALSE, echo=FALSE}
#For each MPRA variant, find the fold change of the allele with the largest absolute fold change
# This creates a column full of NAs for the loop to fill
mpra$max_abs_fc <- NA
# This loop calculates the column of max_abs_fc. It takes the two alleles A and B and calculates which one's absolute value is larger. That larger value is the value of the new column (can be positive or negative). 
for(i in 1:nrow(mpra)){
  if(abs(mpra[i,]$A.log2FC)>abs(mpra[i,]$B.log2FC)){
    mpra[i,]$max_abs_fc<-mpra[i,]$A.log2FC
  }else{
    mpra[i,]$max_abs_fc<-mpra[i,]$B.log2FC
  }
}
```

This runs a loop to find the max fold change p values.
```{r, eval=FALSE, echo=FALSE}
#For each MPRA variant, find the fold change of the allele with the largest absolute fold change
# This creates a column full of NAs for the loop to fill
mpra$max.logPadj_BH<-NA
# This loop calculates the column of max_abs_fc. It takes the two alleles A and B and calculates which one's absolute value is larger. That larger value is the value of the new column (can be positive or negative). 
for(i in 1:nrow(mpra)){
  if(abs(mpra[i,]$A.logPadj_BH)>abs(mpra[i,]$B.logPadj_BH)){
    mpra[i,]$max.logPadj_BH<-mpra[i,]$A.logPadj_BH
  }else{
    mpra[i,]$max.logPadj_BH<-mpra[i,]$B.logPadj_BH
  }
}
```

```{r, eval=FALSE, echo=FALSE}
#For each MPRA variant, find the pics for all the diseases and make max_pics the maximum of all of those values
mpra.pics <- subset(mpra, select=c(MS_pics,RA_pics,UC_pics,T1D_pics,IBD_pics,Crohns_pics,Psoriasis_pics))

# Make the NAs for pics -1
mpra.pics$MS_pics[is.na(mpra.pics$MS_pics)]<- (-1)
mpra.pics$RA_pics[is.na(mpra.pics$RA_pics)]<- (-1)
mpra.pics$UC_pics[is.na(mpra.pics$UC_pics)]<- (-1)
mpra.pics$IBD_pics[is.na(mpra.pics$IBD_pics)]<- (-1)
mpra.pics$T1D_pics[is.na(mpra.pics$T1D_pics)]<- (-1)
mpra.pics$Crohns_pics[is.na(mpra.pics$Crohns_pics)]<- (-1)
mpra.pics$Psoriasis_pics[is.na(mpra.pics$Psoriasis_pics)]<- (-1)
colnames(mpra.pics)
mpra.pics$win_pics <- colnames(mpra.pics)[apply(mpra.pics,1,which.max)] # Make a column that is called win_pics which will give me the name of the pics column with the best 

mpra.pics$max_pics <- NA # This creates a column full of NAs for the loop to fill

for(i in 1:nrow(mpra.pics)){
  if(mpra.pics[i,]$win_pics=="MS_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$MS_pics
  } # if win_pics says MS_pics then the max_pics is MS_pics
  if(mpra.pics[i,]$win_pics=="RA_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$RA_pics
  } # if win_pics says RA_pics then the max_pics is RA_pics
  if(mpra.pics[i,]$win_pics=="UC_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$UC_pics
  } # if win_pics says UC_pics then the max_pics is UC_pics
  if(mpra.pics[i,]$win_pics=="IBD_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$IBD_pics
  } # if win_pics says IBD_pics then the max_pics is IBD_pics
  if(mpra.pics[i,]$win_pics=="Crohns_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$Crohns_pics
  } # if win_pics says Crohns_pics then the max_pics is Crohns_pics
  if(mpra.pics[i,]$win_pics=="Psoriasis_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$Psoriasis_pics
  } # if win_pics says Psoriasis_pics then the max_pics is Psoriasis_pics
   if(mpra.pics[i,]$win_pics=="T1D_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$T1D_pics
  } # if win_pics says T1D_pics then the max_pics is T1D_pics
}

# Make the MPRA pics columns which are NA return to NA
mpra.pics$MS_pics[mpra.pics$MS_pics<0]<- NA
mpra.pics$RA_pics[mpra.pics$RA_pics<0]<- NA
mpra.pics$UC_pics[mpra.pics$UC_pics<0]<- NA
mpra.pics$IBD_pics[mpra.pics$IBD_pics<0]<- NA
mpra.pics$T1D_pics[mpra.pics$T1D_pics<0]<- NA
mpra.pics$Crohns_pics[mpra.pics$Crohns_pics<0]<- NA
mpra.pics$Psoriasis_pics[mpra.pics$Psoriasis_pics<0]<- NA

```

Add max_pics and subset the mpra data to only high pics scores
```{r, eval=FALSE, echo=FALSE}
mpra$max_pics <- mpra.pics$max_pics
mpra <- subset(mpra, max_pics>=0.5)
```

Loop to give zeros and ones based on fc values
```{r, eval=FALSE, echo=FALSE}
a <- NULL
dat.precision<- data.frame(abs(mpra$max_abs_fc)) # create a data frame with max.logPadj_BH and max_abs_fc 
a[1:20] <- data.frame(rep.int(0, 173)) # create a data frame of zeros
dat.precision2 <- data.frame(dat.precision,a) # add the max.logPadj_BH and max_abs_fc columns
names(dat.precision2) <- c("fc",paste0("fc_", seq(0.1, 2, by = 0.1))) # give names to the columns
dat.precision2 <- dat.precision2[,-22]

dat.precision2 <- dat.precision2 |>
  mutate(row = row_number()) |> # not necessary if `fc` values guaranteed unique
  pivot_longer(-c(row, fc)) |>
  mutate(value = 1 * (parse_number(name) <= fc)) |>
  pivot_wider(names_from = name, values_from = value)
```

Loop to give zeros and ones based on padj values
```{r, eval=FALSE, echo=FALSE}
dat.precision3<- data.frame(mpra$max.logPadj_BH) # create a data frame with max.logPadj_BH 
a <- c(0,0,0)
a[1:20] <- data.frame(rep.int(0, 173)) # create b data frame of zeros
dat.precision3 <- data.frame(dat.precision3,a) # add the max.logPadj_BH and max_abs_fc columns
names(dat.precision3) <- c("padj",paste0("padj_", seq(0.1, 2, by = 0.1))) # give names to the columns

dat.precision3 <- dat.precision3 |>
  mutate(row = row_number()) |> # not necessary if `padj` values guaranteed unique
  pivot_longer(-c(row, padj)) |>
  mutate(value = 1 * (parse_number(name) <= padj)) |>
  pivot_wider(names_from = name, values_from = value)
```

Merge the two data frames
```{r, eval=FALSE, echo=FALSE}
dat.precision4 <- merge(dat.precision2, dat.precision3, by="row")
dat.precision4 <- dat.precision4[,-1]
```

Calculating the precision values and then creating a heat map
```{r, eval=FALSE, echo=FALSE}
df <- data.frame(cbind(
c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.1), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.1), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.1),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.1)),

c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.2), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.2), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.2),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.2)),

c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.3), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.3), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.3),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.3)),

c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.4), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.4), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.4),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.4)),

c(Metrics::precision(dat.precision4$padj_0.1, dat.precision4$fc_0.5), 
Metrics::precision(dat.precision4$padj_0.2, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.3, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.4, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.5, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.6, dat.precision4$fc_0.5), 
Metrics::precision(dat.precision4$padj_0.7, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.8, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_0.9, dat.precision4$fc_0.5),
Metrics::precision(dat.precision4$padj_1, dat.precision4$fc_0.5))
))

colnames(df) <- c("FC 0.1","FC 0.2","FC 0.3", "FC 0.4", "FC 0.5")
rownames(df) <- c("Padj_BH 0.1", "Padj_BH 0.2", "Padj_BH 0.3", "Padj_BH 0.4", "Padj_BH 0.5", "Padj_BH 0.6", "Padj_BH 0.7", "Padj_BH 0.8", "Padj_BH 0.9", "Padj_BH 1")

df2 <- df %>%
  rownames_to_column() %>%
  gather(colname, value, -rowname)
head(df2)

ggplot(df2, aes(x = rowname, y = colname, fill = value)) +
  geom_tile() + geom_text(aes(label = round(value, digit=2))) + xlab("Max log p-value Benjamini-Hochberg adjustment") + ylab("Max absolute value fold change") + ggtitle("Precision heatmap for Fold Change") + 
  scale_fill_gradient(low = "gold", high = "firebrick2") +
  coord_fixed()
```



########################

Better loops but not quite correct

########################


Loading the MPRA data (this is the August 25th primary T cell using the T cell filter)
```{r}
mpra.dir <- "/nfs/jray/screens/Mouri_et_al_MPRA/Mouri_et_al_replication/Mouri_replication_rmds/20230825_stimulated_primary_Tcell_glm_replication/john_filter/data/20230825_mpra_merge_Tcell_glm_john.txt"
mpra <- read.table(mpra.dir, sep="\t", header=T)
```

This runs a loop to find the max fold changes.
```{r}
#For each MPRA variant, find the fold change of the allele with the largest absolute fold change
# This creates a column full of NAs for the loop to fill
mpra$max_abs_fc <- NA
# This loop calculates the column of max_abs_fc. It takes the two alleles A and B and calculates which one's absolute value is larger. That larger value is the value of the new column (can be positive or negative). 
for(i in 1:nrow(mpra)){
  if(abs(mpra[i,]$A.log2FC)>abs(mpra[i,]$B.log2FC)){
    mpra[i,]$max_abs_fc<-mpra[i,]$A.log2FC
  }else{
    mpra[i,]$max_abs_fc<-mpra[i,]$B.log2FC
  }
}
```

This runs a loop to find the max fold change p values with the Benjamini Hochberg correction.
```{r}
#For each MPRA variant, find the fold change of the allele with the largest absolute fold change
# This creates a column full of NAs for the loop to fill
mpra$max.logPadj_BH<-NA
# This loop calculates the column of max_abs_fc. It takes the two alleles A and B and calculates which one's absolute value is larger. That larger value is the value of the new column (can be positive or negative). 
for(i in 1:nrow(mpra)){
  if(abs(mpra[i,]$A.logPadj_BH)>abs(mpra[i,]$B.logPadj_BH)){
    mpra[i,]$max.logPadj_BH<-mpra[i,]$A.logPadj_BH
  }else{
    mpra[i,]$max.logPadj_BH<-mpra[i,]$B.logPadj_BH
  }
}
```

This runs a loop to find the max fold change p values with the bonferroni correction.
```{r}
#For each MPRA variant, find the fold change of the allele with the largest absolute fold change
# This creates a column full of NAs for the loop to fill
mpra$max.logPadj_BF<-NA
# This loop calculates the column of max_abs_fc. It takes the two alleles A and B and calculates which one's absolute value is larger. That larger value is the value of the new column (can be positive or negative). 
for(i in 1:nrow(mpra)){
  if(abs(mpra[i,]$A.logPadj_BF)>abs(mpra[i,]$B.logPadj_BF)){
    mpra[i,]$max.logPadj_BF<-mpra[i,]$A.logPadj_BF
  }else{
    mpra[i,]$max.logPadj_BF<-mpra[i,]$B.logPadj_BF
  }
}
```


```{r}
#For each MPRA variant, find the pics for all the diseases and make max_pics the maximum of all of those values
mpra.pics <- subset(mpra, select=c(MS_pics,RA_pics,UC_pics,T1D_pics,IBD_pics,Crohns_pics,Psoriasis_pics))

# Make the NAs for pics -1
mpra.pics$MS_pics[is.na(mpra.pics$MS_pics)]<- (-1)
mpra.pics$RA_pics[is.na(mpra.pics$RA_pics)]<- (-1)
mpra.pics$UC_pics[is.na(mpra.pics$UC_pics)]<- (-1)
mpra.pics$IBD_pics[is.na(mpra.pics$IBD_pics)]<- (-1)
mpra.pics$T1D_pics[is.na(mpra.pics$T1D_pics)]<- (-1)
mpra.pics$Crohns_pics[is.na(mpra.pics$Crohns_pics)]<- (-1)
mpra.pics$Psoriasis_pics[is.na(mpra.pics$Psoriasis_pics)]<- (-1)
colnames(mpra.pics)
mpra.pics$win_pics <- colnames(mpra.pics)[apply(mpra.pics,1,which.max)] # Make a column that is called win_pics which will give me the name of the pics column with the best 

mpra.pics$max_pics <- NA # This creates a column full of NAs for the loop to fill

for(i in 1:nrow(mpra.pics)){
  if(mpra.pics[i,]$win_pics=="MS_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$MS_pics
  } # if win_pics says MS_pics then the max_pics is MS_pics
  if(mpra.pics[i,]$win_pics=="RA_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$RA_pics
  } # if win_pics says RA_pics then the max_pics is RA_pics
  if(mpra.pics[i,]$win_pics=="UC_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$UC_pics
  } # if win_pics says UC_pics then the max_pics is UC_pics
  if(mpra.pics[i,]$win_pics=="IBD_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$IBD_pics
  } # if win_pics says IBD_pics then the max_pics is IBD_pics
  if(mpra.pics[i,]$win_pics=="Crohns_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$Crohns_pics
  } # if win_pics says Crohns_pics then the max_pics is Crohns_pics
  if(mpra.pics[i,]$win_pics=="Psoriasis_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$Psoriasis_pics
  } # if win_pics says Psoriasis_pics then the max_pics is Psoriasis_pics
   if(mpra.pics[i,]$win_pics=="T1D_pics"){
    mpra.pics[i,]$max_pics<-mpra.pics[i,]$T1D_pics
  } # if win_pics says T1D_pics then the max_pics is T1D_pics
}

# Make the MPRA pics columns which are NA return to NA
mpra.pics$MS_pics[mpra.pics$MS_pics<0]<- NA
mpra.pics$RA_pics[mpra.pics$RA_pics<0]<- NA
mpra.pics$UC_pics[mpra.pics$UC_pics<0]<- NA
mpra.pics$IBD_pics[mpra.pics$IBD_pics<0]<- NA
mpra.pics$T1D_pics[mpra.pics$T1D_pics<0]<- NA
mpra.pics$Crohns_pics[mpra.pics$Crohns_pics<0]<- NA
mpra.pics$Psoriasis_pics[mpra.pics$Psoriasis_pics<0]<- NA

```

```{r}
mpra$max_pics <- mpra.pics$max_pics
#mpra <- subset(mpra, max_pics>=0.5)
```

John's heat map. This is wrong. Why?
```{r}
skew<-c(0, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2, 3)
FDR<-c(0.001, 0.01, 0.05, 0.1, 0.15, 0.2, 0.25, 0.5, 0.75, 1)
output3<-NULL
for(s in skew){
  for(f in FDR){
    cur_results<- mpra[,c("rsid","Skew.logFDR","LogSkew","max_abs_fc","max.logPadj_BH","max_pics")] 
    cur_results$Skew_FDR<-10^-cur_results$Skew.logFDR # undo log transform of FDR
    cur_results$Skew<-2^(cur_results$LogSkew) # Undo log transform of Skew
    cur_results2<-cur_results %>% filter(Skew.logFDR<f) # filter results to be less than the FDR threshold
    nrow_skew_emVar<-nrow(cur_results2 %>% filter(Skew>=s & max_pics>=0.1)) # Filter for nrow_skew_emVar is to be greater than the FDR and skew threshold and greater than the high pics threshold
    non_pics<-cur_results %>% filter(max_pics<=0.001) # Filter non_pics is to be less than the low pics threshold
    sampled_non_pics<-non_pics[sample(nrow(non_pics), size=nrow_skew_emVar),] # What is happening here?
    nrow_Skew_non_emVar<-nrow(sampled_non_pics%>% filter(Skew>=s))
    all_top_PICS<-nrow(cur_results %>% filter(max_pics>=0.1))
    precision<-nrow_skew_emVar/(nrow_Skew_non_emVar+nrow_skew_emVar)
    recall<-nrow_skew_emVar/all_top_PICS
    output3<-rbind(output3, data.frame(f, s, precision, recall, nrow_skew_emVar, nrow_Skew_non_emVar, all_top_PICS))
  }
}

ggplot(output3, aes(x=as.character(f), y=as.character(s), fill=precision)) + geom_tile() +  ggtitle("Precision_fixed_negative") + xlab("Skew FDR") + ylab("Skew") + geom_text(aes(label=(round(precision, 2))), color="white", size=4) + 
  scale_fill_gradient(low = "gold", high = "firebrick2")
# ggsave("230425_precision_plot_pics0.9.pdf") # why do I got na's e?
ggplot(output3, aes(x=as.character(f), y=as.character(s), fill=recall)) + geom_tile() +  ggtitle("Recall") + xlab("Skew FDR") + ylab("Skew") + geom_text(aes(label=(round(recall, 2))), color="black", size=4) + 
  scale_fill_gradient(low = "gold", high = "firebrick2") 
# ggsave("230425_recall_plot_pics0.9.pdf")
```

Max's FC heat map. This is wrong. Why?
```{r}
fc_seq <-c(0, 0.25, 0.5, 0.75, 1, 1.25, 1.5, 1.75, 2, 3)
FDR_seq <-c(0.001, 0.01, 0.05, 0.1, 0.15, 0.2, 0.25, 0.5, 0.75, 1)
#FDR_seq <- seq(0.1, 2, by = 0.1)
#fc_seq <- seq(0.1, 2, by = 0.25)
precision.dat <- NULL
for( i in FDR_seq){
  for(j in fc_seq){
   mpra_subset <- subset(mpra, select=c(Skew.logFDR,LogSkew,max_abs_fc,max.logPadj_BH,max_pics))
   FDR_subset <- subset(mpra_subset, Skew.logFDR >= i)
   fc_subset <- subset(FDR_subset, max_abs_fc > j)
   false_positive <- nrow(subset(fc_subset, max_pics<=0.001))
   true_postive <- nrow(subset(fc_subset, max_pics>=0.5))
   precision <- true_postive/(true_postive+false_positive)
   precision.dat <- rbind(precision.dat, data.frame(i, j, precision, true_postive,false_positive))
  }
}
names(precision.dat) <- c("FDR", "FC", "precision", "emVar", "non_Emvar")

ggplot(precision.dat, aes(x = as.character(FDR), y = as.character(FC), fill = precision)) +
  geom_tile() + geom_text(aes(label = round(precision, digit=2))) + xlab("Log Skew FDR p-value") + ylab("Max absolute value fold change") + ggtitle("Precision heatmap for Fold Change") + 
  scale_fill_gradient(low = "gold", high = "firebrick2") 
```